Conservation decision-making under long transients

### 2019-08-13 slides/solarized.Rmd
still going
slides and symlinks

---
title: "Conservation decision-making under long transients"
author: "Carl Boettiger"
institute: "UC Berkeley"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["default","solarized-light.css"]
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: "solarized-light"
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"

@@

## UC Berkeley
---
layout: true
background-color: #fdf6e3
class: center, top

@@

@@

```{r arima_data}
arima_forecast %>% filter(t < 2000) %>%
  ggplot(aes(t)) + 
  geom_line(aes(y=x),  lwd = 1, col=pal[1]) +
  xlim(0,3000) + ylim(0, 1.6)+ ylab("x")

@@

]
.pull-right[
<img src="img/royalsociety/ballons.jpg" width="400", style="float:center">
]

@@

@@

@@

@@

![](img/royalsociety/reef-shark.jpg)

@@

```{r posteriors, fig.width=13, fig.height=6}
label <- c("no switch", "switch", "no ghost")
df <- c("../data/no_switches.csv.xz", "../data/switches.csv.xz", "../data/weak_ghost.csv.xz") %>%
  map_dfr(read_csv, .id = "timeseries") %>% 
  mutate(timeseries = label[as.integer(timeseries)])

@@

p1 <- df %>% 
  ggplot((aes(value, fill=timeseries))) + 
  geom_histogram(alpha=0.9, show.legend = FALSE) + 
  geom_vline(aes(xintercept=0.27, col = "no switch"), col= pal[1], lwd=3) + 
  geom_vline(aes(xintercept=0.28, col = "no ghost"), col = pal[2], lwd=3)

@@

---
background-image: url(img/royalsociety/gentoo-penguin.jpg)
background-position: left
background-size: 140%
class: center, top

@@

## Social

### Utility fn

## Ecological

### Transition fn

@@

# Example systems

@@

---
class: center, middle

# Example Contexts

@@

<img src="img/royalsociety/waiting-in-the-shallows.jpg" width="500">

@@

```{r}
sim <- readRDS("../notebook/sim-ghost-learning.Rds")
states <- seq(0,2, length=140)
actions <- states
Tmax <- 100

@@

```{r}
sim2 <- readRDS("../notebook/sim-ghost-learning-2.Rds")
states <- seq(0,2, length=200)
actions <- states
Tmax <- 100

@@

# An emergent precautionary principle

@@

---
class: center, top
background-image: url(img/royalsociety/bohemian-waxwing.jpg)
background-size: 140%

@@

>>>>

---
title: "Conservation decision-making under long transients"
author: "Carl Boettiger"
institute: "UC Berkeley"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["default", "solarized-light.css"]
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: "solarized-light"
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"

@@

## UC Berkeley
---
layout: true
background-color: #fdf6e3
class: top

@@

---
class: center

@@

```{r arima_data}
arima_forecast %>% filter(t < 2000) %>%
  ggplot(aes(t)) + 
  geom_line(aes(y=x),  lwd = 2, col=pal[1]) +
  xlim(0,3000) + ylim(0, 1.6)+ ylab("x")

@@

- sufficient data

- accurate measurements

- appropriate model

- numerical/computational issues

- ...

.pull-right[
<img src="img/royalsociety/ballons.jpg" width="400", style="float:center">
]

@@

Knowing the correct model is not enough

@@

---
class: top, left

@@

background-image: url(img/royalsociety/reef-shark.jpg)
background-position: center
background-size: 120%
class: center, top

@@

@@

```{r posteriors, fig.width=13, fig.height=6}
label <- c("no switch", "switch", "no ghost")
discrete_posteriors <- c("../data/no_switches.csv.xz", "../data/switches.csv.xz", "../data/weak_ghost.csv.xz") %>%
  map_dfr(read_csv, .id = "timeseries") %>% 
  mutate(timeseries = label[as.integer(timeseries)])

@@

p1 <- discrete_posteriors %>% 
  ggplot((aes(value, fill=timeseries))) + 
  geom_histogram(alpha=0.9, show.legend = FALSE) + 
  geom_vline(aes(xintercept=0.27, col = "no switch"), col= pal[1], lwd=3) + 
  geom_vline(aes(xintercept=0.28, col = "no ghost"), col = pal[2], lwd=3)

@@

---
background-image: url(img/chess.jpg)
background-position: left
background-size: 120%
class: center, bottom

@@

## Societal Dynamics

Utility function $U(x_t,a_t)$

Given state $x_t$ and action $a_t$, balance
costs and benefits.

<img src="img/royalsociety/balance.jpg" width="300px">

@@

.pull-right[

## Ecological Dynamics

Transition function (forecast probabilities) $f(x,a)$

<img src="solarized_files/figure-html/posteriors-1.png" width="400px">

---
class: middle, center

# Potential contexts

*Tipping point or ghost?*

@@

@@

<img src="img/royalsociety/waiting-in-the-shallows.jpg" width="450">

*Reinforcement Learning*

(Markov Decision Process over possible models with Bayesian updating)

@@

```{r learning}
sim <- readRDS("../notebook/sim-ghost-learning.Rds")
states <- seq(0,2, length=140)
actions <- states
Tmax <- 100

@@

```{r less-learning}
sim2 <- readRDS("../notebook/sim-ghost-learning-2.Rds")
states <- seq(0,2, length=200)
actions <- states
Tmax <- 100

@@

# An emergent precautionary principle?

@@

---
class: top, left

# Conclusions

<img src="solarized_files/figure-html/arima_forecast-1.png" width="100px" height="60px"> Ghost attractors look like real attractor but behave differently

<img src="solarized_files/figure-html/wtf-1.png" width="100px" height="60px"> Stochastic ghosts $\neq$ deterministic ones

<img src="solarized_files/figure-html/posteriors-1.png" width="100px" height="60px"> This makes model inference hard!

<img src="img/royalsociety/forest-fire.jpg" width="100px" height="60px"> .. and can lead to very bad decisions!

<img src="solarized_files/figure-html/learning-1.png" width="100px" height="60px"> Reinforcement learning can help deal with uncertainty

<img src="solarized_files/figure-html/less-learning-1.png" width="100px" height="60px"> Precaution is sometimes better than knowledge

---
class: center, top
background-image: url(img/royalsociety/penguin-swim.jpg)
background-size: 140%

@@

Code, slides, etc: <https://github.com/cboettig/decisions-vs-transients>

# Traces for single time-series

```{r}
discrete_posteriors %>% 
  ggplot(aes(t, value, group=chain, col=chain)) + 
  geom_line(show.legend = FALSE) +
  facet_wrap(~timeseries)
```

# Traces for the ensemble fit

```{r}
samples3 %>% 
  ggplot(aes(t, value, group=chain, col=chain)) + 
  geom_line(show.legend = FALSE) 
```
